PoVaTiM: A Possibilistic Valid-Time Model \\ \vspace*{0.04truein}

### 2012-05-24 ijcis-pos-model/sources/preliminaries.tex
last update

A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention. This possibility and necessity is noted here as:
\begin{align}
\text{Possibility that $p$ = $T$ (p holds):} &\\
\nonumber
Pos(p) = \pi_p(T) \label{propholdsposs} \\
\text{Necessity that $p$ = $T$ (p holds):} & \\
\nonumber
Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
\end{align}

Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

This work will deal with ill-known intervals. These are ill-known sets, defined and represented via a start and end point, which will be ill-known values. The elements of the set are the values between the start and end point. A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.

@@

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:triangular-dist}Trapezoidal membership function.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).

A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}. It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. The corresponding membership function definition for this $\mu_T$ is then given by:

@@

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:trapezoidal-dist}Trapezoidal membership function.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}
%\def\JPicScale{0.5}
%\begin{figure}[h!]
%  \centering
%  \input{./graphs/trapezoidal.tex}
%  \caption{Trapezoidal membership function}
%  \label{fig:trapezoidal}
%\end{figure}

@@

So far, we have shown how it can be verified whether or not a set satisfies an ill-known constraint. However, from the examples at the beginning of this section, it is observed that Boolean combinations of constraints are required. For example, the problem of interval evaluation as explained earlier requires that all elements of an interval $[a,b]$ are larger than a value $X$ and at the same time smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied. To allow Boolean combinations of constraints, the following definitions are introduced.
\begin{definition}
\label{def:set-evaluation}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An evaluation function is defined by:
\begin{equation}
\lambda:\Pow(U)\rightarrow\mathbb{B}:A\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
\end{equation}
\end{definition}
Definition \ref{def:set-evaluation} presents the definition of an evaluation function that evaluates a Boolean combination of some basic constraints. Informally, it states that a set $A$ passes the evaluation made by $\lambda$ if the Boolean combination of some propositions equals $T$. This crisp definition can be generalized to the case of ill-known constraints.
\begin{definition}
\label{def:ill-known-sets}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. The uncertainty about the evaluation of a set $A$ by an evaluation function $\lambda$ is then given by:
\begin{equation}
\forall A\in\Pow(U):\pi_{\lambda(A)}=\widetilde{\bool}\Big(\pi_{C_1(A)},...,\pi_{C_n(A)}\Big)\\
\end{equation}
Hereby, $\widetilde{\bool}$ is the possibilistic extension of $\bool$.
\end{definition}
It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only the logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation. Therefore, only the case of Boolean conjunction, Boolean disjunction and Boolean negation will be treated within the scope of this paper. By applying the possibilistic extensions of $\wedge$, $\vee$ and $\neg$, concrete equations are obtained for the calculations of uncertainty about the evaluation of a set by means of an evaluation function $\lambda$. In the case of conjunction (i.e., $\bool=\wedge$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:conjunctive1}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\min_{i=1}^n\Pos\left(C_i(A)\right)\\
\label{eq:conjunctive2}
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\min_{i=1}^n\Nec\left(C_i(A)\right).
\end{eqnarray}
In the case of disjunction (i.e. $\bool=\vee$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:disjunctive}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\max_{i=1}^n\Pos\left(C_i(A)\right)\\
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\max_{i=1}^n\Nec\left(C_i(A)\right).
\end{eqnarray}
Note that by using the functions $\min$ and $\max$ here, there is an implicit assumption that the possibilistic variables $\pi_{C_i}$ are mutual $\min$-dependent in the sense of De Cooman (i.e. non-interactive). For an extensive reading on (in)dependency of possibilistic variables, the reader is referred to~\cite{GertDeCooman1997b},\cite{GertDeCooman1997a},\cite{GertDeCooman1997}. In case of $\neg$, we get:
\begin{eqnarray}
\label{eq:negation}
\forall A\in\Pow(U):\Pos(\neg\lambda(A))&=\\
\nonumber
1-\Nec(\lambda(A))\\
\forall A\in\Pow(U):\Nec(\neg\lambda(A))&=\\
\nonumber
1-\Pos(\lambda(A)).
\end{eqnarray}

@@

\begin{example}
Consider the ill-known values $X = \left[12, 2, 2\right]$ and $Y = \left[18, 2, 1 \right]$. The ill-known value $Z$ is defined using the convex combination $CC$ of constraints $C_1$ and $C_2$:
\begin{align}
\nonumber
CC \triangleq \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
\nonumber
Z \triangleq CC
\end{align}
$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$
Figure \ref{fig:example-ill-known-by-const} illustrates the relations among the variables $X$, $Y$ and $Z$.
\end{example}

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/ill-known-by-constraints.pdf}
\label{fig:example-ill-known-by-const}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{Ill-known values $X$ and $Y$. The grey area represents the ill-known value $Z$ defined by the convex combination of the two ill-known constraints $C_1$ and $C_2$.}
\vspace*{13pt}

>>>>

A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention. This possibility and necessity is noted here as:
\begin{align}
\label{propholdsposs}
\text{Possibility that $p$ = $T$ (p holds):} &\\
\nonumber
Pos(p) = \pi_p(T)  \\
\label{propholdsnecc}
\text{Necessity that $p$ = $T$ (p holds):} & \\
\nonumber
Nec(p) = 1-\pi_p(F) 
\end{align}

Here, equation \eqref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, equation \eqref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

This work will deal with ill-known intervals. These are ill-known sets, defined and represented via a start and end point, which will be ill-known values. The elements of the set are the values between the starting and ending points. A closed ill-known interval with starting point defined by possibilistic variable $X$ and ending point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.

@@

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:triangular-dist}Example of fuzzy number.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. 
%In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).

A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}). It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. Four real values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. The corresponding membership function definition for this $\mu_T$ is then given by:

@@

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:trapezoidal-dist}Example of fuzzy interval.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}
%\def\JPicScale{0.5}
%\begin{figure}[h!]
%  \centering
%  \input{./graphs/trapezoidal.tex}
%  \caption{Trapezoidal membership function}
%  \label{fig:trapezoidal}
%\end{figure}

@@

So far, we have shown how it can be verified if a set satisfies or not an ill-known constraint.

\begin{example}

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/example-ill-known.pdf}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X)$. Possibility and necessity measures are shown in grey the upper and lower graphics respectively. }
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

 Consider an ill-known value given by $X = \left[5, 3, 2 \right]$. $C \triangleq (\leq, X)$ is the ill-known constraint. The set $A$ is $A \subseteq \R^+$. Then, the evaluation of the possibility and the neccesity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively. (See Figure \ref{fig:example-ill-known}).

\begin{align}
\Pos(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\sup_{a \leq R}\pi_{X}(w)\right)\\
\Nec(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\inf_{a > w} 1-\pi_{X}(w)\right) 
\end{align}

\end{example}

 However, from the examples at the beginning of this section, it is observed that Boolean combinations of constraints are required. For example, the problem of interval evaluation as explained earlier requires that all elements of an interval $[a,b]$ are larger than a value $X$ and at the same time smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied. To allow Boolean combinations of constraints, the following definitions are introduced.
\begin{definition}
\label{def:set-evaluation}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An evaluation function is defined by:
\begin{equation}
\lambda:\Pow(U)\rightarrow\mathbb{B}:A\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
\end{equation}
\end{definition}
Definition \ref{def:set-evaluation} presents the definition of an evaluation function that evaluates a Boolean combination of some basic constraints. Informally, it states that a set $A$ passes the evaluation made by $\lambda$ if the Boolean combination of some propositions equals $T$. This crisp definition can be generalized to the case of ill-known constraints.
\begin{definition}
\label{def:ill-known-sets}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. The uncertainty about the evaluation of a set $A$ by an evaluation function $\lambda$ is then given by:
\begin{equation}
\forall A\in\Pow(U):\pi_{\lambda(A)}=\widetilde{\bool}\Big(\pi_{C_1(A)},...,\pi_{C_n(A)}\Big)\\
\end{equation}
Hereby, $\widetilde{\bool}$ is the possibilistic extension of $\bool$.
\end{definition}
It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only the logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation. Therefore, only the case of Boolean conjunction, Boolean disjunction and Boolean negation will be treated within the scope of this paper. By applying the possibilistic extensions of $\wedge$, $\vee$ and $\neg$, concrete equations are obtained for the calculations of uncertainty about the evaluation of a set by means of an evaluation function $\lambda$. In the case of conjunction (i.e., $\bool=\wedge$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:conjunctive1}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\min_{i=1}^n\Pos\left(C_i(A)\right)\\
\label{eq:conjunctive2}
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\min_{i=1}^n\Nec\left(C_i(A)\right).
\end{eqnarray}
In the case of disjunction (i.e. $\bool=\vee$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:disjunctive}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\max_{i=1}^n\Pos\left(C_i(A)\right)\\
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\max_{i=1}^n\Nec\left(C_i(A)\right).
\end{eqnarray}
Note that by using the functions $\min$ and $\max$ here, there is an implicit assumption that the possibilistic variables $\pi_{C_i}$ are mutual $\min$-dependent in the sense of De Cooman (i.e. non-interactive). For an extensive reading on (in)dependency of possibilistic variables, the reader is referred to~\cite{GertDeCooman1997b},\cite{GertDeCooman1997a},\cite{GertDeCooman1997}. In case of $\neg$, we get:
\begin{eqnarray}
\label{eq:negation}
\forall A\in\Pow(U):\Pos(\neg\lambda(A))&=\\
\nonumber
1-\Nec(\lambda(A))\\
\forall A\in\Pow(U):\Nec(\neg\lambda(A))&=\\
\nonumber
1-\Pos(\lambda(A)).
\end{eqnarray}

@@

\begin{example}
Consider the ill-known values $X = \left[12, 2, 2\right]$ and $Y = \left[18, 2, 1 \right]$. The ill-known value $Z$ is defined using the convex combination $CC$ of constraints $C_1$ and $C_2$:
\begin{align}
\nonumber
CC \triangleq \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
\nonumber
Z \triangleq CC
\end{align}
$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$.
Figure \ref{fig:example-ill-known-by-const} illustrates the relations among the variables $X$, $Y$ and $Z$.
\end{example}

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/ill-known-by-constraints.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:example-ill-known-by-const}Ill-known values $X$ and $Y$. The grey area represents the ill-known value $Z$ defined by the convex combination of the two ill-known constraints $C_1$ and $C_2$.}
\vspace*{13pt}

### 2012-05-24 ijcis-pos-model/sources/temporalModel.tex
last update

\begin{itemize}
\item [\textbf{Strictly consistent}] It is not possible for two PVPs of the same object to be valid at the same time. In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
\begin{align}
\label{eq:stricly-consistent}
\forall I,J, \forall t \in \T :\\
%\nonumber
%\left(A, I \right) \in R \wedge \\
%\nonumber
%\left(A, J \right) \in R : \\
\nonumber
\pi_{I}(t) + \pi_{J}(t) \leq 1
\end{align}

\item [\textbf{Ill-consistent}] The overlapping of two PVPs of the same object do overlap. It can be distinghised among the following sub-types:
	\begin{itemize}
	\item [\emph{Co-existence}] Two versions of the same object $A$ may exist at the same time.
	\begin{align}
	\label{eq:co-existence}
	\forall I, \exists ! J, \exists t \in \T : \\
	\nonumber
	\pi_{I}(t) + \pi_{J}(t) \geq 1
	\end{align}
	\item [\emph{Weak-consistence}] Several versions of the same object $A$ may exist at the same time.
	\begin{align}
	\label{eq:weak-consistence}
	\forall I, \exists J, \exists t \in \T : \\
	\nonumber
	\pi_{I}(t) + \pi_{J}(t) \geq 1
	\end{align}
	\end{itemize}
\end{itemize}

>>>>

\begin{itemize}
\item 
\begin{definition}
\textbf{Strictly consistent} It is not possible for two PVPs of the same object to be valid at the same time. In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
\begin{align}
\label{eq:stricly-consistent}
\forall I,J, \forall t \in \T :\\
%\nonumber
%\left(A, I \right) \in R \wedge \\
%\nonumber
%\left(A, J \right) \in R : \\
\nonumber
\pi_{I}(t) + \pi_{J}(t) \leq 1
\end{align}
\end{definition}

\item 
\begin{definition}
 \textbf{Ill-consistent} The overlapping of two PVPs of the same object do overlap. It can be distinghised among the following sub-types:
	\begin{itemize}
	\item 
	      \begin{definition}
	      \emph{Co-existence} Two versions of the same object $A$ may exist at the same time.
		      \begin{align}
		      \label{eq:co-existence}
		      \forall I, \exists ! J, \exists t \in \T : \\
		      \nonumber
		      \pi_{I}(t) + \pi_{J}(t) \geq 1
		      \end{align}
	      \end{definition}

	\item 
	      \begin{definition}
	      \emph{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
		      \begin{align}
		      \label{eq:weak-consistence}
		      \forall I, \exists J, \exists t \in \T : \\
		      \nonumber
		      \pi_{I}(t) + \pi_{J}(t) \geq 1
		      \end{align}
	      \end{definition}

	\end{itemize}
\end{definition}

\end{itemize}
